---
title: "Visualizations to Analyze"
author: "Datatepe"
output: html_document
---
```{r,message = FALSE, warning = FALSE}
# Load necessary library
library(readr)

# Read the CSV file
data <- read_csv("ya_veri.csv")

# Display the data as a table
knitr::kable(data)
```

```{r, fig.show="hide", message = FALSE, warning = FALSE}

# Load necessary libraries
library(ggplot2)
library(readr)
library(dplyr)

# Load the datasetyy
data <- read_csv("ya_veri.csv")

# Filter out 2018 data
data <- filter(data, year != 2018)

# Function to plot gender distribution
plot_gender_distribution <- function(department_name) {
  department_data <- filter(data, department == department_name)
  
  # Define y-axis limits based on department
  y_limits <- if (department_name == "Computer_Eng") {
    c(0, 140)
  } else {
    c(0, 110)
  }
  
  ggplot(department_data) +
    geom_bar(aes(x = factor(year), y = capacity, fill = "Men"), stat = "identity", position = "stack") +
    geom_bar(aes(x = factor(year), y = capacity * (percentage_women / 100), fill = "Women"), stat = "identity", position = "stack") +
    geom_line(aes(x = factor(year), y = valedictorian * 11, group = 1), color = "purple", linewidth = 0.75) +
    geom_point(aes(x = factor(year), y = valedictorian * 11), color = "purple", size = 2) +
    geom_text(aes(x = factor(year), y = capacity, label = paste0(percentage_men, "%")), 
              vjust = -0.5, color = "pink", size = 3, fontface = "bold") +
    geom_text(aes(x = factor(year), y = capacity * (percentage_women / 100), label = paste0(percentage_women, "%")), 
              vjust = 1.5, color = "navy", size = 3, fontface = "bold") +
    geom_text(aes(x = factor(year), y = valedictorian * 11, label = valedictorian), 
              color = "purple", size = 3, fontface = "bold", vjust = -1.5, hjust = 0.5) +
    scale_y_continuous(name = "Capacity", limits = y_limits, breaks = seq(y_limits[1], y_limits[2], by = 10),
                       sec.axis = sec_axis(~ . / 11, name = "Valedictorians", breaks = seq(1, 10, by = 1))) +
    labs(title = department_name, x = "Year", fill = "Gender") +
    scale_fill_manual(values = c("Women" = "pink", "Men" = "navy")) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size = 12),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
          axis.text.y = element_text(size = 10),
          axis.text.y.right = element_text(color = "purple", size = 10),
          axis.title.y.right = element_text(color = "purple", size = 12))
}

# List of departments
departments <- unique(data$department)

# Create and print plots in a single column
for (dept in departments) {
  print(plot_gender_distribution(dept))
}

```

### **Brief Overview**

This page provides key insights into Hacettepe University Engineering Faculty data from 2018-2024, focusing on trends like gender distribution, regional representation, and department capacities.

### Key Observations

1. **Department Capacities**  
   High-demand fields like Computer Engineering have larger capacities, while Mining Engineering shows a decline, reflecting shifts in industrial demand.

2. **Gender Distribution**  
   There's a clear gender imbalance, with women being underrepresented in male-dominated fields like Mechanical Engineering. Food Engineering, in contrast, has more female representation.

3. **Valedictorian Representation**  
   Departments with more women tend to have more female valedictorians, indicating a positive link between gender diversity and achievement.

4. **Geographical Representation**  
   Most students come from the İç Anadolu region, likely due to Hacettepe’s proximity. Regions like Doğu Anadolu have the lowest student representation, suggesting socio-economic barriers.

5. **Ranking Trends**  
   Departments such as Computer Engineering consistently rank high, while fields like Mining Engineering show more fluctuations, possibly due to decreasing interest.

6. **Number of Applications**  
   Computer Engineering and Electrical-Electronic Engineering attract the most applications, reflecting their growing industrial relevance, while niche fields like Hydrogeology Engineering see fewer applicants.

### Why This Data Was Selected

This data was chosen to analyze trends in student preferences, departmental strengths, and societal influences on engineering education. It highlights growing interest in tech fields, sustainability, and the ongoing gender imbalance in traditional engineering disciplines.


```{r,message = FALSE, warning = FALSE}
# Load necessary libraries
library(tidyr)
library(dplyr)
library(ggplot2)
library(forcats)
library(readr)
library(stringr)

# Load the dataset
data <- read_csv("ya_veri.csv")

# Exclude 2018 and reshape the data
data_tidy <- data %>%
  filter(year != 2018) %>%
  select(department, year, akdeniz, dogu_anadolu, ege, guneydogu_anadolu, ic_anadolu, karadeniz, marmara) %>%
  pivot_longer(
    cols = c(akdeniz, dogu_anadolu, ege, guneydogu_anadolu, ic_anadolu, karadeniz, marmara),
    names_to = "region",
    values_to = "percentage_contribution"
  ) %>%
  mutate(region = fct_relevel(region, 
    "ic_anadolu", "marmara", "guneydogu_anadolu", "dogu_anadolu", "akdeniz", "ege", "karadeniz")) %>%
  filter(!is.na(percentage_contribution)) %>%
  group_by(department, year) %>%
  mutate(region = fct_reorder(region, -percentage_contribution, .desc = TRUE)) %>%
  ungroup()

# Define the color palette, including all regions
region_colors <- c(
  "ic_anadolu" = "#e8736b",
  "marmara" = "#54a115",
  "guneydogu_anadolu" = "#053eb0",
  "dogu_anadolu" = "#ff9900",  # Ensure correct name here
  "akdeniz" = "#7705b0",
  "karadeniz" = "#b0054a",
  "ege" = "#d1f299"
)

# Department-Based Plot
ggplot(data=data_tidy, aes(x=year, y=percentage_contribution, fill=region)) +
  geom_bar(stat="identity", position="fill") +
  facet_wrap(~department, ncol = 4) +  # Adjust facet layout
  labs(
    title="Yearly Breakdown by Department",
    x="Year",
    y="Percentage Contribution",
    fill="Region"
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = region_colors) +
  theme_minimal(base_size = 10) +
  theme(
    strip.text = element_text(size = 8, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 6),
    axis.text.y = element_text(size = 8),
    axis.title = element_text(size = 9),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  )



```

### **1. General Information About the Chart**

-   The chart presents how different regions contributed to university departments over multiple years.

-   It highlights proportional contributions by region and shows how trends shifted across departments.

-   The visualization uses a **stacked bar chart** to represent percentage contributions, enabling easy comparison of regional dynamics.

### **2. Key Observations**

#### **a. Dominant Regions**

-   **İç Anadolu** has consistently been the leading contributor across most departments, reflecting the university's location.

-   **Marmara** also contributes significantly but shows a gradual decline in recent years.

#### **b. Regional Shifts**

-   **Southeastern Anatolia** experienced a sharp increase in 2023 contributions, likely due to the earthquake, boosting interest in civil and geology engineering.

-   **Aegean** and **Mediterranean** regions display variability, with notable recoveries in recent years.

#### **c. Stable Trends**

-   **Karadeniz** shows steady contributions over time, with minor fluctuations.

### **3. Insights**

-   Proximity plays a key role, as **İç Anadolu** dominates due to accessibility.

-   External factors like the **2023 earthquake** significantly impacted regional contributions, particularly from **Southeastern Anatolia**.

-   Some regions, like **Karadeniz**, show stability, while others, such as **Aegean**, reveal periodic shifts, reflecting local factors.

### **Conclusion**

This analysis highlights regional preferences and shifts in contributions to university departments. It provides a clear understanding of dominant regions, stable contributors, and regions affected by external influences. The visualization effectively communicates these trends, offering valuable insights into regional dynamics.

